A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Rewrite content based similarity sections

\subsection*{Content Based Similarity}
A simple approach is to recommend products that have similar content. Each product
has includes a short description which we can use to construct a textual 
content-based similarity function. To do this, we represent each
product as a vector of tf-idf values for each word in its descriptions. By using
tf-idf, we reduce the relevance of frequent words and boost that of rare ones.
This emphasizes the words that better explain this product versus all the other 
ones in the catalogue. The vector representations are the length of the
vocabulary. They are very sparse as most words do not appear in any given
product description. With this tf-idf represenation, we can define the 
$ContentSimiarity$ functoin as follows:

@@

This similarity function yields the following recommendation algorithm which we
call the $ContentBased Recommender$: For each
product, scan through the products in the opposite catalogue and calculate the
content-based similarity score with each. Predict recommendations between this
product and each of the top k most similar products in the other catalogue.

While this is a quadric-time operation, it can be sped up using matrix
multiplication. Furthermore, the division into different categories reduces the
impact of the quadratic effect. In our case, Macy's entire catalogue consists
of over $60,000$ products, making a quadratic time algorithm very slow.
However, the average category has only one or two thousand products, which is 
significantly more manageable size for quadratic operations.

\subsection*{Leveraging Recommendations}
This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
to capture similarity between products by looking exclusively at the words used
to describe them. One limitation here is that different brands often use
different vocabularies to describe the same thing, such as ``off-white'' versus
``ivory.'' Another issue is that the best recommendation for a product is often
more subtle than simply the most similar other product. For instance, a striped
black-and-white button-up shirt may be an excellent recommendation for a blue
checkered button-up, even though the two have ostensibly different patterns and
colors.

We can alleviate some of these issues by leveraging the recommendations we
already have between products in the same catalogue. One approach is to consider
the neighborhood of each product in their recommender graphs, that is a product
along with each of the products it is recommended with in its own catalogue.
When comparing two products in different catalogues, we use the same tf-idf
similarity function from before to compare each pair of products between the two
neighborhoods. Mathematically, we construct the similarity function
NeighborhoodSimilarity defined as:

NeighborhoodSimilarity definition...

Assuming the recommendations are good within each of the catalogues, this
approach can help both of the issues brought up with the simple content-based
recommender. Since a neighborhood has many products, each describing themselves
with potentially different vocabularies, there is a higher likelihood that the
neighborhood contains multiple synonyms and thus decreases the potential of
missing due to different vocabularies. Furthermore, the recommender graphs
capture some of the user behavior not expressed in product descriptions.
Re-examining the button-up shirts example, the neighborhood of a striped shirt may
include checkered shirts and vise-versa. In this case, the two neighborhoods
would share the words ``striped'' and ``checkered'' and thus the two previously
dissimilar shirts would now be considered similar.

\subsection*{Latent Concepts}
Although the NeighborhoodSimilarity method alleviates some of the issues of
content-similarity, it is still fundamentally based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways than their descriptions. If a dress is elegant and modern, we
want to recommend another dress that is elegant and modern, even if those words
are not explicitly in the descriptions of either. These types of latent concepts
can be intangible and difficult to define. However, if we assume that the given
recommendations are truly based on these underlying concepts, we can attempt to 
learn the latent concepts by analyzing the graphs.

One approach to learning latent concepts is Latent Semantic Indexing or LSI. In
traditional LSI, used for Natural Language Processing, concepts are learned over 
a dataset consists of documents, each containing terms. Singular Value 
Decomposition (SVD) is applied to the term-document matrix and the resulting 
three matrices are truncated to the number of latent concepts desired. The 
resulting left matrix is a term-concept matrix, where each row is a term 
represented in concept space.

We apply the same idea to the product catalogues in order to learn the latent
concepts of each product. Analogous to documents containing terms, we look at
user sessions which contain the products viewed by that user. Although we do not
have actual user data, we can simulate user sessions by traversing the
recommender graphs.

To simulate user sessions, we use a random walk technique on the graph. We
assume that the edges of a recommender graph indeed connect products that a user
would co-view. Thus traversing the edges of a graph is a reasonable simulation
of a user session. For details on these random walks, see the section titled
Random Walks in the Implementation section below.

@@

In all, the LSI approach consists of four phases. First we simulate user
sessions by randomly walking the recommender graphs. Second we learn concept
spaces for products in each catalogue via SVD. Third we find a mapping between
concept spaces. And lastly we find the nearest neighbors of each product in the
other catalogue's concept space. These nearest neighbors are the predicted
recommendations between products in different catalogues.

@@

\subsection*{Random Walks}
The first step of the LSI approach is to simulate user sessions by randomly
walking the recommender graphs. There are a few ways this could be implemented.
One method is to randomly assign a random walker a starting node and then allow
it to randomly walk the edges, with a probability of termination at each step.
Running this many times is a straight-forward model of user sessions.

To speed up the simulation, we calculate a series of product-specific stationary
distributions. As with any random walk technique, we first interpret the 
recommender graph as a transition matrix. In this walk, for each product we create a
product-specific transition matrix which has added teleportation edges back the
that product. We calculate the stationary distributions of each of these
matrices, one per product. We use these product-specific stationary
distributions as our user sessions. Since there is only one user session per 
product, we have many fewer sessions than we would with a real dataset. However,
each of these stationary distributions can be interpreted as an aggregation of
an infinite number of random walkers starting at each node and travelling until
a random termination time, as in the original model.

@@

We imagine that recommender graphs are constructed as follows: A retailer first
builds a similarity matrix, using a collaborative filtering method. In this
matrix, all similarity relationships are symmetric. Next, the retailer modifies
the values of the matrix to include popularity. As a simple model, we say that
the popularity-modified similarity score from Product i to Proudct j is the
simple similarity score at (i, j) multiplied by the popularity of Product j. In
this way, more popular items are more frequently recommended.

We see signs of this in the undirected recommender graphs. Although each product
has at most four outgoing recommendations on macys.com, we see that some
products have over one hundred incoming recommendations. With this in mind, we
approximate the popularity of each product as the logarithm of the degree of
that product in its undirected recommender graph. We choose to model it as the
logarithm because a few of the products tend to have degrees in the hundreds.
Without some way of scaling back these popularities, these few very popular
products tend to take over all recommendations.

@@

Assuming this model of recommender graph construction, which separates similarity
from popularity, we seek to isolate similarity from the recommender graph. Since
popular products were promoted for popularity, we demote them when walking the
graphs to create user sessions. In practice, we divide the likelihood of walking
an edge by the popularity of the product on the other side of that edge. Thus
the concepts learned on that matrix better represent products independent of
popularity.

Proceeding as before, we find the nearest products in concept space. However,
before deeming the nearest neighbors to be the best recommendations, we divide
the distance to each product by its popularity. Since we model distance in 
concept space to be inversely proportional to similarity, this reincorporation
of popularity is analagous to multiplying similarity by popularity, as done in
the original construction of the recommender graph.

A final way in which we incorporate popularity is in the number of
recommendations alotted per product. Since we model the recommender graph as an
undirected graph, different nodes have different degrees. A product that is
popular within its catalogue is likely to be more popular across catalogues as
well. Therefore, we allow popular products to have more recommendations across
the two catalogues, proportional to their popularity.

@@

A more impressive baseline is that of the Simple Content-Based approach, 
outlined earlier in the paper. This recommender uses information exclusively
from the descriptions of the products and ignores the recommender graphs 
entirely. A third, somewhat similar baseline comes from the Neighborhood 
Content-based variant, which looks at the content similarity between 
neighborhoods of products.

@@

The first baseline is in fact an upper bound of sorts on the mapping. This
baseline recommender is called the One Model Recommender. Unlike the other
baselines, this recommender is not a valid one, as it cheats by using
information from the unpartitioned graph. The One Model Recommender learns a
latent concept space over the entire recommender graph by allowing the random
walkers to traverse it before it is partitioned. In every other respect, it is
the same as the LSI Recommender. However, since products from both catalogues
are represented in the same concept space, there is no need for a mapping
between concept spaces.

>>>>

\subsection*{Content Based Similarity}
A simple approach that does not rely on user data is to recommend products that
have similar content. Each product page includes a short description which we 
can use to construct a textual content-based similarity function. To do this, 
we represent each product as a vector of tf-idf values for each word in its 
description. By using tf-idf, we reduce the relevance of frequent words and 
boost that of rare ones. This emphasizes the words that better represent this 
product versus all the other ones in the catalogue. The vector representations 
are the length of the vocabulary. These are sparse vectors as most words do 
not appear in any given product description. With this tf-idf represenation,
we can define the $ContentSimiarity$ functoin as follows:

@@

This similarity function yields the following recommendation algorithm which we
call the $ContentBased Recommender$: For each product, calculate its
$ContentSimilarity$ score with each of the products in the other catalogue.
Predict recommendations between this product and the top 2 most similar products 
in the other catalogue.

While this is a quadric-time operation, it can be executed efficiently using 
matrix multiplication. Furthermore, the division into different categories 
reduces the impact of the quadratic effect. In our case, the entire Macy's 
catalogue consists of over $60,000$ products, making a quadratic time algorithm 
slow. However, the average category has only one to two thousand products, which
is a significantly more manageable size for quadratic operations.

\subsection*{Leveraging Recommendations}
This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
to capture similarity between products by looking exclusively at the words in
their descriptions. One issue here is that brands tend to use different 
vocabularies to describe the their products. One may refer to a color as 
``off-white'' while another calls the same color ``ivory.'' To the $ContentBased
Recommender$, these words are as different as ``black'' and ``white.'' Another 
issue is that the best recommendation for a product may not be the most similar
other product. For instance, a striped black-and-white button-up shirt could be 
an excellent recommendation for a blue checkered button-up, even though the two 
have ostensibly different descriptions.

We can alleviate some of these issues by leveraging the recommendations we
already have between products within the same catalogue. One approach is to 
consider the {\emph neighborhood} of each product, that is that product along 
with each of its neighbors in the recommendation graph. With this we create a
new similarity function, $NeighborhoodSimilarity$. This function uses the 
$ContentSimilarity$ function from before to compare every pair of products 
between the two neighborhoods:

\begin{align}
NeighborhoodSimilarity(P_1, P_2) &= \sum\limits_{S\in
Neigh(P_1)}\sum\limits_{T\in Neigh(P_2)}
ContentSimilarity(S, T) 
\end{align}
where $Neigh(P)$ is the neighborhood of Product P. We use this similarity
function to construct the $Neighborhood Recommender$, which is analgous to the
$ContentBased Recommender$.

Assuming that the edges of the recommendation graphs are in fact good
recommendations, this approach can help both of the issues brought up with the
$ContentBased Recommender$. Since a neighborhood has many products, each 
describing themselves using potentially different vocabularies, there is a 
higher likelihood that the neighborhood contains multiple synonyms for the
underlying concept that describes their similarity. This decreases the
likelihood of false negatives due to different vocabularies. Furthermore,
the recommendation graphs capture some of the user behavior that cannot be
expressed in product descriptions. Re-examining the button-up shirts example, 
the neighborhood of a striped shirt may include checkered shirts and vise-versa.
In this case, the two neighborhoods would share the words ``striped'' and 
``checkered'' and thus the two previously dissimilar shirts would now correctly
be deemed similar.

\subsection*{Latent Concepts}
Although the $Neighborhood Recommender$ alleviates some of the issues of 
content-based similarity, it is still fundamentally based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways. If a dress is elegant and modern, we would like to recommend 
another dress that is elegant and modern, even if those words are not explicitly
in the descriptions of either. These types of underlying concepts can be 
intangible and difficult to define. However, if we assume that the given
recommendations are truly based on such latent concepts, we can attempt to learn 
them by analyzing the graphs.

One common approach to learning latent semantic factors is to apply Singular Value
Decomposition to . In
traditional LSI, used for Natural Language Processing, concepts are learned over 
a dataset consists of documents, each containing terms. Singular Value 
Decomposition (SVD) is applied to the term-document matrix and the resulting 
three matrices are truncated to the number of latent concepts desired. The 
resulting left matrix is a term-concept matrix, where each row is a term 
represented in concept space.

We apply the same idea to the product catalogues in order to learn the latent
concepts of each product. Analogous to documents containing terms, we look at
user sessions which contain the products viewed by that user. Although we do not
have actual user data, we can simulate user sessions by traversing the
recommendation graphs.

To simulate user sessions, we use a random walk technique on the graph. We
assume that the edges of a recommendation graph indeed connect products that a user
would co-view. Thus traversing the edges of a graph is a reasonable simulation
of a user session. For details on these random walks, see the section titled
Random Walks in the Implementation section below.

@@

In all, the LSI approach consists of four phases. First we simulate user
sessions by randomly walking the recommendation graphs. Second we learn concept
spaces for products in each catalogue via SVD. Third we find a mapping between
concept spaces. And lastly we find the nearest neighbors of each product in the
other catalogue's concept space. These nearest neighbors are the predicted
recommendations between products in different catalogues.

@@

\subsection*{Random Walks}
The first step of the LSI approach is to simulate user sessions by randomly
walking the recommendation graphs. There are a few ways this could be implemented.
One method is to randomly assign a random walker a starting node and then allow
it to randomly walk the edges, with a probability of termination at each step.
Running this many times is a straight-forward model of user sessions.

To speed up the simulation, we calculate a series of product-specific stationary
distributions. As with any random walk technique, we first interpret the 
recommendation graph as a transition matrix. In this walk, for each product we create a
product-specific transition matrix which has added teleportation edges back the
that product. We calculate the stationary distributions of each of these
matrices, one per product. We use these product-specific stationary
distributions as our user sessions. Since there is only one user session per 
product, we have many fewer sessions than we would with a real dataset. However,
each of these stationary distributions can be interpreted as an aggregation of
an infinite number of random walkers starting at each node and travelling until
a random termination time, as in the original model.

@@

We imagine that recommendation graphs are constructed as follows: A retailer first
builds a similarity matrix, using a collaborative filtering method. In this
matrix, all similarity relationships are symmetric. Next, the retailer modifies
the values of the matrix to include popularity. As a simple model, we say that
the popularity-modified similarity score from Product i to Proudct j is the
simple similarity score at (i, j) multiplied by the popularity of Product j. In
this way, more popular items are more frequently recommended.

We see signs of this in the undirected recommendation graphs. Although each product
has at most four outgoing recommendations on macys.com, we see that some
products have over one hundred incoming recommendations. With this in mind, we
approximate the popularity of each product as the logarithm of the degree of
that product in its undirected recommendation graph. We choose to model it as the
logarithm because a few of the products tend to have degrees in the hundreds.
Without some way of scaling back these popularities, these few very popular
products tend to take over all recommendations.

@@

Assuming this model of recommendation graph construction, which separates similarity
from popularity, we seek to isolate similarity from the recommendation graph. Since
popular products were promoted for popularity, we demote them when walking the
graphs to create user sessions. In practice, we divide the likelihood of walking
an edge by the popularity of the product on the other side of that edge. Thus
the concepts learned on that matrix better represent products independent of
popularity.

Proceeding as before, we find the nearest products in concept space. However,
before deeming the nearest neighbors to be the best recommendations, we divide
the distance to each product by its popularity. Since we model distance in 
concept space to be inversely proportional to similarity, this reincorporation
of popularity is analagous to multiplying similarity by popularity, as done in
the original construction of the recommendation graph.

A final way in which we incorporate popularity is in the number of
recommendations alotted per product. Since we model the recommendation graph as an
undirected graph, different nodes have different degrees. A product that is
popular within its catalogue is likely to be more popular across catalogues as
well. Therefore, we allow popular products to have more recommendations across
the two catalogues, proportional to their popularity.

@@

A more impressive baseline is that of the Simple Content-Based approach, 
outlined earlier in the paper. This recommender uses information exclusively
from the descriptions of the products and ignores the recommendation graphs 
entirely. A third, somewhat similar baseline comes from the Neighborhood 
Content-based variant, which looks at the content similarity between 
neighborhoods of products.

@@

The first baseline is in fact an upper bound of sorts on the mapping. This
baseline recommender is called the One Model Recommender. Unlike the other
baselines, this recommender is not a valid one, as it cheats by using
information from the unpartitioned graph. The One Model Recommender learns a
latent concept space over the entire recommendation graph by allowing the random
walkers to traverse it before it is partitioned. In every other respect, it is
the same as the LSI Recommender. However, since products from both catalogues
are represented in the same concept space, there is no need for a mapping
between concept spaces.
